How Landcover Errors Impact Our Understanding of Socioeconomic and Environment Change

### 2017-03-20 Rmd/c10-location-impact.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...cropland
crclass <- lapply(cropl, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

@@

# ...carbon
cdat <- c("gti" = gti_c$f1, lc_c$f1)  # combine C 1 km sets
cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
  msk <- x$forest < 99.6
  cmu <- calc(x, mean)
  raster::mask(cmu, msk, maskvalue = 0)
})
cclass <- lapply(cdatf, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

@@

>>>>

### Identify high value areas
```{r, eval = FALSE}
# top decile of...
# ...cropland
crclass <- lapply(cropl, function(x) {  # x <- cropl$sa30
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})

@@

# ...carbon
cdat <- c(gti_c$f1, lc_c$f1)
cdatf <- lapply(cdat, function(x) {  # mask out non-cropland areas and take mu C
  msk <- x$forest < 99.6
  cmu <- calc(x, mean)
  raster::mask(cmu, msk, maskvalue = 0)
})
cclass <- lapply(cdatf, function(x) {  # x <- cyld_agg$f1[[1]]
  r <- x * (x > 0)
  r[r == 0] <- NA
  rq <- quantile(r, c(0, 0.9, 1))
  types <- cut(r, breaks = rq)
  types > 1
})
names(cclass) <- names(pclass)

@@

# average km off
mudvec <- unlist(mudists)
median(mudvec[!like(names(mudvec), "sa30")])
# mean(mudvec)

### 2017-03-20 Rmd/c4-cropland-2.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

### Total error
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
# paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
# gti <- lapply(paths, raster)
# names(gti) <- c("g2007", "g2011")
# gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
# snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

@@

# calculate km2 cropland in LC datasets
awgts <- lapply(area_wgts, function(x) x[[1]])
lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
  gtia <- cellStats(gti_agg[[x]]$g2011 / 100 * awgts[[x]], sum)
  lca <- sapply(snms, function(y) {
    cellStats(lc_agg[[x]][[y]] / 100 * awgts[[x]], sum)
  })
  o <- c(gtia, lca)
}))
colnames(lckm2)[1] <- "gti"
rownames(lckm2) <- lev2

@@

totkm2 <- cellStats(!is.na(gti_agg$f1$g2011) , sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```
<a href="#top">Back to top</a>

@@

# calculate bias/MAE for data
a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
bacc2011 <- rbind(a, b)

>>>>

### Total error
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
fact <- c(5, 10, 25, 50, 100)
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
gti <- lapply(gti, function(x) raster::mask(x, namask))  # apply namask
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

@@

# calculate km2 cropland in LC datasets
awgts <- lapply(area_wgts, function(x) x[[1]])
lckm2 <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
  gtia <- cellStats(gti_agg11[[x]] / 100 * awgts[[x]], sum)
  lca <- sapply(snms, function(y) {
    cellStats(lc_agg[[x]][[y]] / 100 * awgts[[x]], sum)
  })
  o <- c(gtia, lca)
}))
colnames(lckm2)[1] <- "gti"
rownames(lckm2) <- lev

bias_statsw

# mean (stays about the same)
lckmmu <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
  gtia <- cellStats(gti_agg11[[x]] / 100, mean)
  lca <- sapply(snms, function(y) {
    cellStats(lc_agg[[x]][[y]] / 100, mean)
  })
  o <- c(gtia, lca)
}))
colnames(lckmmu)[1] <- "gti"
rownames(lckmmu) <- lev

# variance declines with scale, as expected
lckmvar <- do.call(rbind, lapply(1:length(lev), function(x) {  # x <- 1
  gtia <- cellStats(gti_agg11[[x]] / 100, var)
  lca <- sapply(snms, function(y) {
    cellStats(lc_agg[[x]][[y]] / 100, var)
  })
  o <- c(gtia, lca)
}))
colnames(lckmvar)[1] <- "gti"
rownames(lckmvar) <- lev

@@

totkm2 <- cellStats(!is.na(gti_agg$f1$g2011), sum)
save(stats2007, stats2011, totkm2, refkm2, 
     file = full_path(p_data, "total-cropland.rda"))
```
<a href="#top">Back to top</a>

@@

# gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
# gti_agg11 <- lapply(gti_agg, function(x) x$g2011)

# calculate bias/MAE for data
a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
bacc2011 <- rbind(a, b)
save(bacc2011, file = fp(p_data, "cropland-werr-2011.rda"))

### 2017-03-20 Rmd/c7-et.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

@@

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

@@

caption <- paste("Biases and mean absolute errors (as \\%) for",              
                 "evapotranspiration variables derived from a 29-year time",
                 "series calculated by the VIC model, including the average", 
                 "total ET for the", 
                 "3 months of the year when ET is highest, the annual mean", 
                 "and the minimum and maximum annual ETs in the time series.")
mutab_xtab <- xtable::xtable(mutab, digits = 1, caption = caption)
print(mutab_xtab, type = "latex", file = "paper/figures/et-bias.tex",
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)
```
<a href="#top">Back to top</a>

>>>>

# cropland data
load(full_path(p_data, "MZshapes.Rdata"))  # SA shape
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- raster::mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

@@

mos <- format(ISOdatetime(2000,1:12,1,0,0,0),"%b")
lnames <- sapply(1981:2008, function(x) paste0(mos, "_", substr(x, 3, 4)))
lnames <- lnames[1:length(lnames)]
# length(lnames) == nlayers(etba[[1]])
etba <- lapply(etba, function(x) {
  b <- raster::mask(x, mask25k, maskvalue = 0)
  names(b) <- lnames
  b
})
names(etba) <- anms

@@

caption <- paste("Biases and mean absolute errors (as \\%) for",              
                 "evapotranspiration variables derived from a 29-year time",
                 "series calculated by the VIC model, including the average", 
                 "total ET for the", 
                 "3 months of the year when ET is highest, the annual mean", 
                 "and the minimum and maximum annual ETs in the time series.")
mutab_xtab <- xtable::xtable(mutab, digits = 1, caption = caption)
print(mutab_xtab, type = "latex", file = "paper/figures/et-bias.tex",
      tabular.environment = "longtable", floating = FALSE, 
      caption.placement = "top", include.rownames = FALSE)

et_err <- mutab
save(et_err, file = fp(p_data, "et-err-bias.rda"))
```
<a href="#top">Back to top</a>

### 2017-03-20 Rmd/c8-yield-production.Rmd
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

@@

## Analyses
### Cropland fractions
#### Get provincial $cf$ factors from 1 km GTI dataset
```{r, eval=FALSE}
# rasterize provinces and stack with gti raster, convert to data.table
# provr <- rasterize(prov, y = gti, field = "id_1", 
#                    filename = "external/ext_data/provinces.tif")
provr <- raster("external/ext_data/provinces.tif")
provr <- mask(provr, namask)
gtis <- stack(list("prov" = provr, "f" = gti))

@@

# data.table calculations
gti_dt <- na.omit(as.data.table.raster(gtis, xy = TRUE))
setkey(gti_dt, prov)
cf <- gti_dt[, list("carea" = sum(f), "parea" = length(f), 
                    "cf" = sum(f) / length(f)), by = prov]
# cftest - cf[, 2:4, with = FALSE]  # close, but diff caused by few NAs in provr

# data.table approach is much more efficient

@@

### Plot difference maps
```{r, eval=FALSE}
dnms <- ls()[grep("[y_|p_]diff*.*d", ls())]
stnms <- ls()[grep("stats_", ls())]
disaggl <- lapply(dnms, function(x) get(x))
statsl <- lapply(stnms, function(x) get(x))
rng <- lapply(statsl, function(x) {
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
})

@@

# check error stats - do they match alternate approach?
ref <- lapply(gti_agg, function(x) x$gti)
test_err <- "perr" #"yerr"
evst <- "wma" # "wm"
for(i in c("f1", "f25", "f10")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms,  
                      get(evst), aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst), rweight = FALSE, 
                      aweight = FALSE)[, j, with = FALSE]
    a3 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst))[, j, with = FALSE]
    s <- stack(get(test_err)[[i]][[j]], ref[[i]], awgts[[i]])
    sv <- getValues(s)
    sv <- sv[which(!is.na(rowSums(sv))), ]  # remove NAs
    if(evst == "wm") {
      print(a1 == round(weighted.mean(sv[, 1], sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(sv[, 1]), 2))
      print(a3 == round(weighted.mean(sv[, 1], sv[, 2] * sv[, 3]), 2))
    } else if(evst == "wma"){
      print(a1 == round(weighted.mean(abs(sv[, 1]), sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(abs(sv[, 1])), 2))
      print(a3 == round(weighted.mean(abs(sv[, 1]), sv[, 2] * sv[, 3]), 2))
    }
  }
}

@@

save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
save(cyld_agg, cpagg1, file = full_path(p_data, "yieldprod-1km.rda"))

>>>>

# cropland data
load(full_path(p_data, "d_grid_act.rda"))  # actual diffence grids
paths <- full_path(p_data, dir(p_data, paste0("cover*.*11sum_mask.tif$")))
gti <- raster(paths) / 100  # gti 2011
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
gti <- raster::mask(gti, namask)  # apply mask to GTI
# cellStats(!is.na(gti), sum)

@@

## Analyses
### Cropland fractions
#### Get provincial $cf$ factors from 1 km GTI dataset
```{r, eval=FALSE}
# rasterize provinces and stack with gti raster, convert to data.table
# provr <- rasterize(prov, y = gti, field = "id_1", 
#                    filename = "external/ext_data/provinces.tif")
provr <- raster("external/ext_data/provinces.tif")
provr <- raster::mask(provr, namask)
gtis <- stack(list("prov" = provr, "f" = gti))

@@

# data.table calculations
gti_dt <- na.omit(as.data.table.raster(gtis, xy = TRUE))
setkey(gti_dt, prov)
cf <- gti_dt[, list("carea" = sum(f), "parea" = length(f), 
                    "cf" = sum(f) / length(f)), by = prov]
# cftest - cf[, 2:4, with = FALSE]  # close, but diff caused by few NAs in provr 
# data.table approach is much more efficient
```
<a href="#top">Back to top</a>

@@

### Plot difference maps
```{r, eval=FALSE}
dnms <- ls()[grep("[y_|p_]diff*.*d", ls())]
stnms <- ls()[grep("^stats_", ls())]
disaggl <- lapply(dnms, function(x) get(x))
statsl <- lapply(stnms, function(x) get(x))
rng <- lapply(statsl, function(x) { # x <- statsl[[1]]
  ilims <- c(ceiling(min(sapply(x, function(y) y[3, ]))), 
             floor(max(sapply(x, function(y) y[21, ]))))
  olims <- range(sapply(x, function(y) range(y)))
  c(olims[1], ilims, olims[2])
})

@@

# check error stats - do they match alternate approach?
ref <- lapply(gti_agg, function(x) x$gti)
test_err <- "perr" #"yerr"
evst <- "wma" # "wm"
for(i in c("f1", "f25", "f100")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms,  
                      get(evst), aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst), rweight = FALSE, 
                      aweight = FALSE)[, j, with = FALSE]
    a3 <- bias_statsw(ref[[i]], awgts[[i]], get(test_err)[[i]], snms, 
                      get(evst))[, j, with = FALSE]
    s <- stack(get(test_err)[[i]][[j]], ref[[i]], awgts[[i]])
    sv <- getValues(s)
    sv <- sv[which(!is.na(rowSums(sv))), ]  # remove NAs
    if(evst == "wm") {
      print(a1 == round(weighted.mean(sv[, 1], sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(sv[, 1]), 2))
      print(a3 == round(weighted.mean(sv[, 1], sv[, 2] * sv[, 3]), 2))
    } else if(evst == "wma"){
      print(a1 == round(weighted.mean(abs(sv[, 1]), sv[, 2], na.rm = TRUE), 2))
      print(a2 == round(mean(abs(sv[, 1])), 2))
      print(a3 == round(weighted.mean(abs(sv[, 1]), sv[, 2] * sv[, 3]), 2))
    }
  }
}

@@

save(list = ls(), file = full_path(p_data, "yield-bias.rda"))
save(out_tab, file = fp(p_data, "yield-prod-accbias.rda"))
save(cyld_agg, cpagg1, file = full_path(p_data, "yieldprod-1km.rda"))

### 2017-03-20 paper/SI.tex
Long-overdue commit, including fully revised manuscript, edits to various code files to undertake analyses necessary to make revisions

\setpagewiselinenumbers
%\modulolinenumbers[5]
%\linenumbers
\begin{center}\section*{\Large Supplementary Methods and Results}\end{center}
%\begin{table}[ht]
%\centering
%\caption{Confidence classes for the forcing data based on bias-correction weights.}
%\begin{tabular}{cc}
%  \hline
%  Confidence class & Bias-correction weight \\ 
%  \hline
%High & 0.75-1.00 \\ 
%Medium-High & 0.50-0.75\\ 
%Medium-Low & 0.25-0.50\\ 
% Low & 0.05-0.25 \\ 
% None & $<$0.05 \\ 
%   \hline
%\end{tabular}
%\end{table}  
\section{\large Overview}
This document provides supplementary information on methods and results from the different components of the analysis described in the main text.  The code used to conduct the analysis is presented after sections 1-X.

@@

We also calculated how different methods of calculating the bias and accuracy statistics impacted out findings, using three different methods. First, we simply took the straight averages across the entire country, which substantially understates bias and accuracy because cropland covers only 10\% of the country, and all landcover products successfully discriminate the non-cropped regions.  We also compared the average error metrics extracted from within just the agricultural regions of the maps, defined here as the union of areas having $>$0.5\% cropland in the reference and each test map. This also tends to understate bias, because the area having $<$5\% cropland is much larger than areas having higher densities of cropland (see Fig. 4 frequencies of cells per five percentile bins of cropland density). Finally, we calculated a density-independent metrics, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%). The results of these three alternate metrics are shown in Figure 3.

>>>>

\setpagewiselinenumbers
%\modulolinenumbers[5]
%\linenumbers
\begin{center}\section*{\Large Supplementary Methods and Results}\end{center}
%\begin{table}[ht]
%\centering
%\caption{Confidence classes for the forcing data based on bias-correction weights.}
%\begin{tabular}{cc}
%  \hline
%  Confidence class & Bias-correction weight \\ 
%  \hline
%High & 0.75-1.00 \\ 
%Medium-High & 0.50-0.75\\ 
%Medium-Low & 0.25-0.50\\ 
% Low & 0.05-0.25 \\ 
% None & $<$0.05 \\ 
%   \hline
%\end{tabular}
%\end{table}  
\section{\large Overview}
This document provides supplementary information on methods and results from the different components of the analysis described in the main text.  The code used to conduct the analysis is presented after section 7.

@@

We also calculated how different methods of calculating the bias and accuracy statistics impacted out findings, using three different methods. First, we simply took the straight averages across the entire country, which substantially understates bias and accuracy because cropland covers only 10\% of the country, and all landcover products successfully discriminate the non-cropped regions.  We also compared the average error metrics extracted from within just the agricultural regions of the maps, defined here as the union of areas having $>$0.5\% cropland in the reference and each test map. This also tends to understate bias, because the area having $<$5\% cropland is much larger than areas having higher densities of cropland (see Fig. 4 frequencies of cells per five percentile bins of cropland density). Finally, we calculated a density-independent metric, which is similar to the density weighted mean calculated for our main analysis, which was calculated by averaging the mean bias and MAE values within different levels of cropland cover (divided in increments of 5\%). The results of these three alternate metrics are shown in Figure 3.
